Von Neumann-Morgenstern and Savage Theorems for Causal Decision Making

### 2020-08-30 Theory_and_decision/Draft/Theory_and_decision_draft.tex
cambios en intro

These criteria are the basis for many techniques used in Artificial Intelligence; as previously mentioned, RL algorithms learn \textit{optimal policies} in such a way that any action prescribed by such optimal policy achieves the maximum expected utility \citep{sutton1998reinforcement,webb2007game}. In general, any algorithm which relies on the von Neumann-Morgenstern or Savage Theorems, such as the optimal policies from Reinforcement Learning, is based on associative relations which are expressed using correlations or probability distributions.

It is a natural question how to formalize rational decision making when causal information is present; answering such question is relevant given the importance of causal relations in everyday life. Given that human beings actually use causal information while making choices and the importance of the associative decision making results, it is desirable to have an explicit, computationally implementable criterion for decision making analogous to those by von Neumann-Morgenstern and Savage.

The previous question has been already considered by \cite{nozick1969newcomb}, \cite{lewis1981causal} and by \cite{joyce1999foundations} without an explicit optimality criterion for decision making and also by \cite{pearl2009causality}, who provides an optimality criterion for decision making under causal-controlled uncertainty when the causal mechanism which controls the environment is known by the decision maker.

>>>>

These criteria are the basis for many techniques used in Artificial Intelligence; as previously mentioned, RL algorithms attempt to learn \textit{optimal policies} in such a way that the actions prescribed by such optimal policy achieve the maximum expected utility by satisfying the Bellman Equations \citep{sutton1998reinforcement,webb2007game}. In general, any algorithm which relies on the von Neumann-Morgenstern or Savage Theorems, such as the optimal policies from Reinforcement Learning, is based on \textit{associative} relations which are expressed using patterns found in the data in terms of correlations encoded in probability distributions.

It is a natural question \textit{how to formalize rational decision making when causal information is present}; answering such question is relevant given the importance of causal relations in everyday life as well as in science. Given that human beings actually use causal information while making choices and the importance of the associative decision making results, it is desirable to have an explicit, computationally implementable criterion for decision making analogous to those by von Neumann-Morgenstern and Savage in order to provide the foundations for Causal Reinforcement Learning.

The previous question has been already considered by \cite{nozick1969newcomb,lewis1981causal} and by \cite{joyce1999foundations} without an explicit optimality criterion for decision making as well as by \cite{pearl2009causality}, who provides an optimality criterion for decision making under causal-controlled uncertainty \textit{when the causal mechanism which controls the environment is known} by the decision maker.
